Iterative Reduction: Parsing and Interpreting Code Without Trees

### 2025-01-21 Paper/paper.tex
fix error

\section{Introduction}
\yoni{comment}
Nowadays, most programming language interpreters follow the following pipeline (see \cite{dragon})
\begin{enumerate}
    \item Lexing: lexically analyze the input code, separating it into atomic \textit{tokens}.
    \item Parsing: convert the parsed tokens into a \textit{parse tree}: a tree which represents the code according to the language's grammar.
\yoni{comment}
    \item Interpreting: recursively descend the tree and execute the code according to the instructions embedded in the tree.
\end{enumerate}
For example, we can parse the following expression:

@@

We propose a novel method for the process of interpretation, which doesn't use parse trees.
But our algorithm is not limited to this, and can be altered to produce parse trees as well (see \S\ref{producing-parse-trees}).
This means that our algorithm can be utilized as part of the pipeline for compilers and traditional interpreters.

The idea of the algorithm is as follows: the input is a stream of tokens, which are indivisible characters in any language.
Given a string of tokens $\mb{\sigma^1}\cdots\mb{\sigma^n}$ (for readability, tokens are written in boxes), we form rules for combining two tokens.
For example, $\mb{\ttt1}$ and $\mb{\ttt+}$ can be combined together to give a token $\mb{\ttt{1,+}}$.
\yoni{comment}
Then if we follow a token of the form $\mb{\ttt{x,+}}$ with a token of the form $\mb{\ttt y}$, the tokens are combined to form a token $\mb{\ttt{x+y}}$.
So for example:
\[ \mbt{1} \mbt{+} \mbt{2} \mbt{+} \mbt{3} \to \mbt{1,+} \mbt2 \mbt+ \mbt3 \to \mbt{3} \mbt+ \mbt3 \to \mbt{3,+} \mbt3 \to \mbt6 \]
But what about \ttt{1+2*3}?
We'd get
\[ \mbt1 \mbt+ \mbt2 \mbt* \mbt3 \to \mbt{1,+} \mbt2 \mbt* \mbt3 \to \mbt3 \mbt* \mbt3 \to \mbt{3,*} \mbt3 \to \mbt9 \]

>>>>

\section{Introduction}
Nowadays, most programming language interpreters follow the following pipeline (see \cite{dragon})
\begin{enumerate}
    \item Lexing: lexically analyze the input code, separating it into atomic \textit{tokens}.
    \item Parsing: convert the parsed tokens into a \textit{parse tree}: a tree which represents the code according to the language's grammar.
    \item Interpreting: recursively descend the tree and execute the code according to the instructions embedded in the tree.
\end{enumerate}
For example, we can parse the following expression:

@@

We propose a novel method for the process of interpretation, which doesn't use parse trees.
But our algorithm is not limited to this, and can be altered to produce parse trees as well (see $\S$\ref{producing-parse-trees}).
This means that our algorithm can be utilized as part of the pipeline for compilers and traditional interpreters.

The idea of the algorithm is as follows: the input is a stream of tokens, which are indivisible characters in any language.
Given a string of tokens $\mb{\sigma^1}\cdots\mb{\sigma^n}$ (for readability, tokens are written in boxes), we form rules for combining two tokens.
For example, $\mb{\ttt1}$ and $\mb{\ttt+}$ can be combined together to give a token $\mb{\ttt{1,+}}$.
Then if we follow a token of the form $\mb{\ttt{x,+}}$ with a token of the form $\mb{\ttt y}$, the tokens are combined to form a token $\mb{\ttt{x+y}}$.
So for example:
\[ \mbt{1} \mbt{+} \mbt{2} \mbt{+} \mbt{3} \to \mbt{1,+} \mbt2 \mbt+ \mbt3 \to \mbt{3} \mbt+ \mbt3 \to \mbt{3,+} \mbt3 \to \mbt6 \]
But what about \ttt{1+2*3}?
We'd get
\[ \mbt1 \mbt+ \mbt2 \mbt* \mbt3 \to \mbt{1,+} \mbt2 \mbt* \mbt3 \to \mbt3 \mbt* \mbt3 \to \mbt{3,*} \mbt3 \to \mbt9 \]
